---
title: "Exploration of Methods for Geographic Visualization"
author: "Tova Ibbotson"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Static Visualization
```{r load packages}
library(knitr)
library(ggplot2)
library(dplyr)
require(maps)
require(viridis)

theme_set(theme_void())

```

```{r load data, warning=FALSE}
isolates <- read.csv("isolates2.csv")

# Restrict to isolates that have a state identified
isolate_states <- isolates %>% 
  filter(!state == "Not Specified") %>% 
  rename(cdc_region = region) %>% 
  rename(region = state) %>%
  filter(!(region == "Other"|region == "PR")) %>%
  filter(Collection.year==c(2015:2022))

isolate_states$region <- tolower(state.name[match(isolate_states$region,state.abb)])

```

```{r missing, results='hide'}
sum(is.na(isolates$Lat.Lon))/nrow(isolates)
sum(!isolates$state=="Not Specified")/nrow(isolates)

```

Over 98% of the data in the latitude-longitude variable are missing, so geographic visualization of the data must be done at the state-level. We have state-level location information for roughly 60% of the isolates, so geographic analysis will be restricted to the 1060 observations collected between 2015 and 2022 with state-level location data. Figure 1 shows the number of isolates found in each state over the time period.


```{r map state}
states_map <- map_data("state")

print_map_states <- 
  isolate_states %>% 
  group_by(region) %>% 
  summarize(n_isolates = n()) %>%
  left_join(states_map, by = "region") %>%
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n_isolates), color = "white") +
  scale_fill_viridis(option = "D", name = "Number of Isolates") +
  borders("state") +
  labs(title = "Figure 1", subtitle = "Number of Isolates by State")

print_map_states

# Create table of count of isolates by state
a <- isolate_states %>% 
  group_by(region) %>% 
  summarize(n_isolates = n())

```

```{r map cdc region}
# Create table of count of isolates by state
b <- isolate_states %>% 
  group_by(cdc_region) %>% 
  summarize(n_isolates = n())

print_map_cdc <- 
  left_join(b, isolate_states, by = "cdc_region") %>%
  left_join(states_map, by = "region") %>%
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n_isolates), color = "white") +
  scale_fill_viridis(option = "D", name = "Number of Isolates") +
  borders("state") +
  labs(title = "Figure 2", subtitle = "Number of Isolates by Microbiological Monitoring Region")

print_map_cdc

```

-create a map of BIFSCo regions, see if we can add borders around groups of states

```{r bifsco map}

bifsco_states <- isolates %>%
  distinct(state,region) %>%
  filter(!(state == "Not Specified" | state == "Other" |
             state == "HI" | state == "AK" | state == "PR")) %>%
  rename(cdc_region = region) %>%
  rename(region = state)

bifsco_states$region <- tolower(state.name[match(bifsco_states$region,state.abb)])

print_bifsco_map <- bifsco_states %>%
  left_join(states_map, by = "region") %>%
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = factor(cdc_region)), color = "white") +
  geom_sf(fill = "transparent", color = "gray20", size = 1, 
          data = . %>% group_by(cdc_region) %>% summarise()) +
  scale_fill_discrete(name = "Region") +
  borders("state") +
  labs(title = "Figure 3", subtitle = "Map of CDC Microbiological Monitoring Regions")

print_bifsco_map

bifsco_states %>% group_by(cdc_region) %>% summarise()



```


-next, facet wrap by a variable with <5 levels

## Interactive Visualization

shiny with above facet wrap as filter, if time allows try to add hover summary box






\newpage
# References


# Appendix: All code for this report

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

```
